\chapter{Conclusions}

In this thesis we reproduced a very interesting and elegant approach developed by Botero et al. \cite{alonso} to analyse the three body problem in the plane related to the Hall effect. We showed that in the classical case, this problem is integrable due to the decoupling of the guiding centres and the linear momenta in the regimen of big magnetic fields. Moreover, the canonical transformation of relative coordinates can be manipulated to codify the shape of the triangle in a Bloch sphere. With this analysis, the study of the orientation and shape of the triangle can be studied naturally in terms of the dynamical and geometrical angular velocity of the system, which are in fact the angular variables in the angle-action formalism.\\

In the quantum counterpart of the problem, a similar analysis was shown, where the symmetry of the system has a more noticeable role as this mentioned relative coordinates transformation implements a Schwinger quantum angular momentum $SU(2)$. Moreover, the reduction of the degrees of freedom of the problem due to de big magnetic field can be compared to the quantum Hall effect on the plane, where in this regime, the electron states are restricted to the base state.\\

The elegance of the study by Botero et al. \cite{alonso}, and its graceful correlation with the quantum counterpart, leaded us to look for a similar formalism to describe the three-body problem on the spherical surface. The fact that Haldanes \cite{haldane} formalism of the quantum base states for Hall effect on the sphere is very similar to that of the planar quantum Hall effect, leaded us to believe that a formalism similar to Botero et al. on the regime of big magnetic fields could be easily developed for the sphere. Moreover, as we analysed the one-body problem on the sphere as an introduction to the three-body formalism, we discovered that in terms of trajectories, there were many similarities that indicated that the finding of the wanted analogue analysis was actually possible.\\

When we tried to develop the mentioned formalism on the sphere, we found that although there are some analogues for some quantities and some coincidences on the trajectories, the structure of the algebra of guiding center transformation on the sphere is very different from the plane, impeding then further simplification of the problem trough the crucial relative coordinate transformation.\\

We studied the three-body problem on the sphere in Cartesian, spherical and stereographic coordinates, all of which have their advantages and flaws. We found that if we want a very similar analysis from that of the plane, the stereographic projection, due to its nature and its results, is the most suitable candidate. Otherwise, the mapping of the plane formalism to the sphere is not trivial nor even that similar.  This fact is supported by the fact that on the Haldane formalism, some spinor representation introduced to describe the base states is deeply related to the stereographic projection.\\

Given all these hints, we believe that a proper development of a spherical formalism similar to the planar from Botero et al. \cite{alonso}, can be performed through the use of stereographic coordinates. However, as this mapping demonstrated to be not trivial at all, we leave this study for further work.\\

